---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# peeky  <picture><source media="(prefers-color-scheme: dark)" srcset="man/figures/logo-peeky-dark-animated.svg"><img src="man/figures/logo-peeky-light-animated.svg" align="right" alt="Logo: cartoon eyes peeking over an app card whose corner peels back to reveal source code." width="150"/></picture>

<!-- badges: start -->
[![R-CMD-check](https://github.com/coatless-rpkg/peeky/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coatless-rpkg/peeky/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `peeky` package helps you extract, examine, and run the source code from Shiny 
applications that have been converted to run in the browser using [Shinylive][slexplain].
It works with both standalone applications and [Quarto][quarto] documents 
containing Shinylive components through the [quarto-shinylive][qsl] extension,
supporting both R and Python Shiny applications.

## About Shinylive

[Shinylive][slexplain] converts existing Shiny applications
to run entirely in the web browser using [WebAssembly][wasm] versions of
R ([webR][webr]) and Python ([Pyodide][pyodide]), eliminating the 
need for a computational server. This means all 
application files are accessible to users by design as they are downloaded
to the user's browser when the application is loaded.

The Shinylive ecosystem consists of four main components: 

- Shinylive Web Assets: [posit-dev/shinylive](https://github.com/posit-dev/shinylive)
- R Shiny App Converter Package: [posit-dev/r-shinylive](https://github.com/posit-dev/r-shinylive)
- Python Shiny App Convert Package: [posit-dev/py-shinylive](https://github.com/posit-dev/py-shinylive)
- Shinylive Quarto Extension: [quarto-ext/shinylive](https://github.com/quarto-ext/shinylive)

## There Are No Secrets in Shinylive

The `peeky` package was developed to demonstrate a fundamental truth about 
Shinylive applications stressed by its developers: **"There are no secrets."**

Unlike traditional Shiny applications where server-side code remains private,
Shinylive apps run entirely in the web browser, making **all** associated 
files accessible to users. This includes the source code, data, and any other
files used by the application. As a result, Shinylive applications are
transparent by design.

This package was developed as part of ongoing discussions in STATS 290 about
Shiny application security, transparency, and deployment options. It serves as
a practical demonstration of the differences between traditional server-side 
applications and modern browser-based alternatives.

## Installation

You can install the development version of peeky from [GitHub](https://github.com/) with:

```r
# install.packages("remotes")
remotes::install_github("coatless-rpkg/peeky")
```

## Usage

To use the package, load it into your R session:

```r
library(peeky)
```

Inside the package, there are three main functions:

| Function | Description |
|----------|-------------|
| `peek_shinylive_app()` | Universal function that handles both standalone apps and Quarto docs |
| `peek_standalone_shinylive_app()` | Specifically for standalone Shinylive applications |
| `peek_quarto_shinylive_app()` | Specifically for Quarto documents with Shinylive components |


### Extracting Shinylive Applications

We suggest using the `peek_shinylive_app()` function as it can handle both 
standalone Shinylive applications and Quarto documents with Shinylive components.
For instance, if we take the main Shinylive extension website, we get:

```{r}
#| label: peek-shinylive-app
#| cache: true
# Choose where the files are written (here, a temporary directory)
out_dir <- file.path(tempdir(), "shinylive-apps")
peeky::peek_shinylive_app("https://quarto-ext.github.io/shinylive/", output_dir = out_dir)
```

This would be equivalent to if we ran the following:

```r
peeky::peek_quarto_shinylive_app(
  "https://quarto-ext.github.io/shinylive/",
  output_path = out_dir
)
```

The output location is a required argument, so the package never writes to your
working directory unless you ask it to: a relative path such as `"my-apps"` is
created under the current working directory, while an absolute path is used
as-is. Each application is placed in a subdirectory named `app_1`, `app_2`,
etc. We can also set the output format to `quarto` to extract the files into a
single Quarto document.

```{r}
#| label: peek-shinylive-app-output-dir
#| cache: true

# Extract the applications into a single Quarto document
peeky::peek_quarto_shinylive_app(
  "https://quarto-ext.github.io/shinylive/",
  output_format = "quarto",
  output_path = file.path(tempdir(), "shinylive-apps.qmd")
)
```

We can switch to the `peek_standalone_shinylive_app()` function if we know that
the URL is a standalone Shinylive application. For example, if we take the
example application used in the conversion tutorial from [an app.R to an R Shinylive app](https://github.com/coatless-tutorials/convert-shiny-app-r-shinylive)
on GitHub, we get:

```{r}
#| label: peek-shinylive-standalone-app
#| cache: true
peeky::peek_standalone_shinylive_app(
  "https://tutorials.thecoatlessprofessor.com/convert-shiny-app-r-shinylive/",
  output_dir = file.path(tempdir(), "standalone-app")
)
```

## License

AGPL (>= 3)

## Evolution

This package represents a more refined and comprehensive approach compared to 
our [earlier tutorial][peeksl]
that focused solely on standalone R Shinylive applications. 

## Ethical Considerations

This package is for educational purposes. Users should:

- Use responsibly
- Respect intellectual property rights
- Understand that viewable code doesn't imply permission to reuse
- Consider this when designing their own applications

## Acknowledgements

Thanks to the Shinylive, [webR](https://docs.r-wasm.org/webr/latest/) and [Pyodide](https://pyodide.org/en/stable/) teams for enabling browser-based data science.

[quarto]: https://quarto.org
[slexplain]: https://shiny.posit.co/py/get-started/shinylive.html
[qsl]: https://github.com/quarto-ext/shinylive
[webr]: https://docs.r-wasm.org/webr/latest/
[pyodide]: https://pyodide.org/en/stable/
[peeksl]: https://github.com/coatless-tutorials/peeking-at-an-r-shinylive-app-source-code
[wasm]: https://webassembly.org/
